---
title: "Classification Models Part A Tutorial"
output:
  html_document:
    df_print: paged
---
#1. Load the park dataset as above
```{r}
library(tidyverse)
library(ISLR)
park <- read_csv('http://www.dsi.uminho.pt/~pcortez/forestfires/forestfires.csv')
```
#2. Create a new variable called fire such that fire = 1 if area > 0 and fire=0 otherwise
```{r}
for (i in 1:length(park$area)){
  if (park$area[i] > 0.0){
    park$fire[i] <- 1}
    else{
      park$fire[i] <- 0
    }
}
```
#3. Convert the following variables to factor type, to tell glm these are categorical:
#-  X, Y, month, day
#-  ideally using a mutate_at statement
```{r}
head(park)
park%>%
  mutate(X=as.factor(X),Y=as.factor(Y),month=as.factor(month),day=as.factor(day))
```
#4. Display the number of observations in each fire category
```{r}
sum(park$fire > 0)
sum(park$fire == 0)
```
#5. Run a logistic model on the outcome fire using all other variables as regressors except area
```{r}
mod <- lm(fire ~ X + Y + month + day + FFMC + DMC + DC + ISI + temp + RH + wind + rain, data = park)
summary(mod)
```
#6. Make a scatter plot of the estimated probabilities of fire as a function of temp, with different colors for month
```{r}
park%>%
  ggplot(aes(x = temp, y = fire, color = month)) + geom_jitter(alpha = .5)
```

