STOR 455 Homework #1

High Peaks
```{r}
library(readr)
library(mosaic)
library(Stat2Data)
data(HighPeaks)
data(Perch)
data(SpeciesArea)
```
A) Least Squares Line: y = 0.6843x + 2.0482
```{r}
lm(formula=Time~Length, data=HighPeaks)
View(HighPeaks)
```
B) Scatterplot
```{r}
mod1=lm(Time~Length, data=HighPeaks)
plot(Time~Length, data=HighPeaks)
abline(mod1)
```
C) Time and residual model predicted for climbing Panther Peak
Time: 14.12788
Residual: -0.591394241
```{r}
mod1$residuals
2.0842+(0.6843)*(17.6)
```
D) Largest positive residual: 4.003444 Seward
```{r}
max(mod1$residuals)
```

E) Most negative residual: -2.449137 Mt. Colden
```{r}
min(mod1$residuals)
```
F) Conditions
```{r}
plot(mod1$residuals~mod1$fitted.values)
abline(a=0,b=0)
```
Linearity - yes, the data seems to be following a linear pattern
Zero Mean - yes, the errors are centered at zero
```{r}
mod1=lm(Time~Length, data=HighPeaks)
plot(Time~Length, data=HighPeaks)
abline(mod1)
```
Constant Variance - no, there seems to be more points toward the middle and right
```{r}
hist(mod1$residuals)
```
```{r}
qqnorm(mod1$residuals)
qqline(mod1$residuals) 
```
Independence - The histogram looks good, as it is slightly skewed and the Q-Q plot also looks good as most of the points are on or are close to the line.

Perch
A)Linear model
y = 3.163x-5.079
```{r}
mod1=lm(Length~Weight, data=Perch)
plot(Length~Weight, data=Perch)
abline(mod1)
```
B)
```{r}
plot(log(Length)~log(Weight), data=Perch)
betterModel = lm(log(Length)~log(Weight), data=Perch)
abline(betterModel)
summary(betterModel)
```
C) Checking linearity/conditions
```{r}
plot(betterModel$residuals~betterModel$fitted.values)
abline(a=0,b=0)
```
There are many more points to the right of the residual plot.
```{r}
hist(betterModel$residuals)
```
This histogram of the residuals appears to be skewed to the left.
```{r}
qqnorm(betterModel$residuals)
qqline(betterModel$residuals) 
lm(formula= log(Weight)~log(Length), data=Perch)
```
The Q-Q plot looks good as most of the points are close to the line.

D)
Length = 19 cm
Weight = 55.018