Universities are typically opaque, bureaucratic institutions. Many public schools such as the University of North Carolina system make [data about university employees publically](http://www.newsobserver.com/news/databases/public-salaries/) available. Using this dataset what can we learn about the university?


## Instructions

- Submit responses to Sakai
- **Submit only the html output from your markdown response.** Files that do not *knit* will **not be considered valid.**

If there is a knitting error, and you can't identify the source after substantial effort, change the `eval` option in the problematic chunk to `FALSE` so that you can at least turn in the rest of the assignment. 

Done like so
```
{r, eval = FALSE}

```

**But you will not get any credit for that particular response.** All responses must display the output and code requested in order to receive credit.

In addition,

- use the `tidyverse` and `dplyr` functions to answer the following questions
- submit just the .Rmd file to Sakai
- pay careful attention to the questions
- **return** something means just print it out. Similarly for questions **What is ...** just print out the answer

Example

Q: What is the sum of the first 10 positive integers?
```{r}
sum(1:10)
```

- **create** or **save** means save a variable

Create a variable `x` with the sum of the first 10 positive integers.
```{r}
x <- sum(1:10)
```



# The data

```{r}
library(tidyverse)
```


First download the data set from Ryan Thornburg's website (it is also on stor390's github repo). You may want to set `cache=TRUE` for this chunk so you don't have to download the data set every time you run the script.
```{r setup, message=FALSE, cache=FALSE}
data <- read_csv("http://ryanthornburg.com/wp-content/uploads/2015/05/UNC_Salares_NandO_2015-05-06.csv")
```

Take a first look at the data
```{r}
str(data)
head(data)
```

## Tips

- Read through the `dplyr` lecture slides to date.

- `fte` means full time employee


# Assignment

## Question 0
The data come from the same source as this [N\&O database](https://www.newsobserver.com/news/databases/public-salaries/article11863496.html), but captured in May 2015.

Based on the information in the link above and on inspection of the dataset using `head`, `str` etc., briefly answer the following questions. If you do not have clear evidence for an answer, say so and justify a guess. If you know nothing at all from the resources above, say 'I don't know'.

You need only use the resources above to answer these questions.

**Q:** Who generates the data in this dataset?

**A:**

**Q:** Does the dataset include employee pay from *all* sources?

**A:**

**Q:** Does the `totalsal` variable give salary actually paid to the employee in a given year?

**A:**

**Q:** What does the `employed` variable represent?

**A:**

## Question 1
Return a data frame with columns: name, dept, age,totalsal 
```{r}
#
```

Return a data frame with columns: statesal, nonstsal,totalsal using one of the `select_helpers` functions
```{r}
#
```

## Question 2 
Rename the `fte` column to `fulltime`. Make sure this change is saved (i.e. `data <- ...`).
```{r}
#
```

## Question 3 

What is the mean salary in the Neurosurgery department?

```{r}
#
```


Return a data frame with employee's in the Neurosurgery department making more than $500,000.  
```{r}
#
```


**Q:** Why might these professors be so well paid?

**A:**

## Question 4 
What is the total amount that full time Dermatology employees get paid?


```{r}
#
```


## Question 5

Create a data frame called radio_dept whose rows are the employees from the Radiology department. 

- include only the following columns: name, position, age, nonstsal, totalsal. 
- order the employees by salary

Use pipes `%>%` to make your code readable.

```{r }
#
```


## Question 6

Create a data frame called `dept_summary`  whose rows are the departments and  whose columns are: department size, mean department salary, median department salary, and maximum salary (using totalsal for salary). 

```{r }
#
```

Order the departments by highest mean salary and print the 10 highest paid departments. 


```{r }
#
```

Order the departments by highest median salary and print the 10 highest paid departments. 

```{r }
#
```

**Q:** Why do these lists differ? If you were asked for the top 10 best paid departments at UNC which summary would you choose and why?

**A:**

## Question 7
How many departments have at least 10 employees?
```{r }
#
```


## Question 8
Which department hired the most people in 2010? Hint: you probably have to modify `hiredate`.
```{r }
#
```

## Question 9
Make a new data frame called `dept_big` that includes only departments with at least 20 employees, and save it as a new object (i.e. `dept_big <- ...`)
```{r }
#
```


## Question 10
Repeat the analysis of Q 6 but with `dept_big` instead of the original dataset, and using `dept_summary_big` instead of `dept_summary`. Answer all questions of Q 6 but for `dept_summary_big`, including the short-answer response.
