```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
```

## Instructions
Answer each of the questions given below. Do so in a markdown file with html output.

**Submit only the HTML file.** 

Remember to `knit` often and use the techniques we discussed in class to handle code chunks with errors you cannot fix. Homework submitted in a format other than the html output from your document will not be accepted.

## Scoring
All questions worth the same amount, **three points.**

## A Secret Santa sampler
The Secret Santa Christmas tradition is one in which each person is randomly assigned a single other person who they must give a gift. More typical in the U.S. holiday season is that each person gives each other person a gift. Secret Santa therefore is an economical way to enjoy the holiday gift-giving tradition.

There are many [variants of the game](https://en.wikipedia.org/wiki/Secret_Santa), and versions for other occasions.

Various websites will generate Secret Santa matchups for a given list of participants.

**You will create a Secret Santa sampler yourself in this homework!**

### Rules for your program

- Input: Vector of participant names
- Output: Assignment of gift givers to gift receivers for all participants
- Rules for assignments:
  - each participant gives only one gift
  - each participant receives only one gift
  - participants cannot give themselves gifts

### Homework structure
As often is the case, you might see a different, perhaps more efficient, way to create this program. For example, you could create a program that deals with data frames, not lists.

However, I break the task into steps to split up points awarded and to guide those without coding experience. 

The tasks are meant to give students practice with stuff seen recently in class, and with useful tools we did not have time to cover, particularly in the [purrr package](https://purrr.tidyverse.org/reference/index.html).


### Program outline

We will use the following terminology

- a **santa** will refer to the person giving the gift
- a **giftee** will refer to the person receiving the gift

Our program will work as follows

1. Given a vector of participant names, create a list of vectors each of which has two elements, `santa` and `giftee`, representing one `santa, giftee` pair---excluding cases where santa and giftee are the same person

2. Repeat the following until no possible `santa`, `giftee` choices remain, ie. till the output from part b is empty
  a. select one `santa`, `giftee` at random
  b. update the list of possible pairs to ensure the `santa` chosen in part a can no longer be chosen as a santa, and `giftee` from part a can no longer be chosen as a giftee
  c. return the output from parts a and b

### Example of program step 2

Say our list of participants is as follows

```
c('spike', 'virgil', 'diana', 'little bear')
```

Then the list of possible pairs would look like this


```
[[1]]
   santa   giftee 
 "spike" "virgil" 

[[2]]
   santa   giftee 
"virgil"  "diana" 

[[3]]
        santa        giftee 
      "diana" "little bear" 

[[4]]
        santa        giftee 
     "virgil" "little bear" 

[[5]]
        santa        giftee 
"little bear"       "diana" 

[[6]]
        santa        giftee 
"little bear"      "virgil" 

[[7]]
   santa   giftee 
 "diana" "virgil" 

[[8]]
        santa        giftee 
"little bear"       "spike" 

[[9]]
        santa        giftee 
      "spike" "little bear" 

[[10]]
  santa  giftee 
"spike" "diana" 

[[11]]
  santa  giftee 
"diana" "spike" 

[[12]]
   santa   giftee 
"virgil"  "spike" 
```

If in part 2a we select

```
  santa   giftee 
"virgil"  "spike" 
```

then the result from part 2b will look like this

```
[[1]]
   santa   giftee 
 "spike" "virgil" 

[[2]]
   santa   giftee 
 "diana" "virgil" 

[[3]]
        santa        giftee 
"little bear"      "virgil" 

[[4]]
  santa  giftee 
"spike" "diana" 

[[5]]
        santa        giftee 
"little bear"       "diana" 

[[6]]
        santa        giftee 
      "spike" "little bear" 

[[7]]
        santa        giftee 
      "diana" "little bear" 
```

All cases where `virgil` is santa OR `spike` is giftee have been removed.


# Q1

Run the following code to **create a vector of names**, with which we will test pieces of our program.

```{r}
library(tidyverse)
names <- c('spike', 'virgil', 'diana', 'little bear')
```

Use the `cross` or `cross2` [functions in the tidyverse](https://purrr.tidyverse.org/reference/cross.html) to **create a list of all possible pairs**, including illegal ones where santas and giftees are the same. 

**Display the result**

```{r}
data <- list(santa = names,
  giftee = names)

allPairs<- data %>%
  cross() %>%
  map(.,unlist)

allPairs
```

### Hints:

- create a list with two elements, `santa = names` and `giftee = names`
- use `cross` on this list
- the output from `cross` is not exactly what we want, since it returns a list of lists
- convert the output of `cross` from a list of lists to a list of vectors by applying `unlist` to each element
- you will want to use `map` or `lapply` for the previous step

**Your output at this point will include illegal santa giftee pairs**, e.g.
```
  santa   giftee 
"virgil"  "virgil" 
```

# Q2
We need to modify your response to Q1 to remove pairs where santa and giftee are the same person.

**We will do this in two different ways** across Q2, 3, 4 for the sake of learning some coding tools.

**Write a function** that takes a single argument, a vector of length two, which will be a single one of the pairs in your list output from Q1.

**Your function should return `TRUE` or `FALSE`** based on whether the first and second values of the input are the same or different, respectively.

For example, if the input is 

  santa   giftee 
"virgil"  "virgil" 

your function should return `TRUE`.


```{r}
my_function <- function(x){
  if (x[1] == x[2])
    TRUE
  else 
    FALSE
}
allPairs[[1]]
my_function(allPairs[[1]])

allPairs[[2]]
my_function(allPairs[[2]])
```

# Q3
Remember that `filter` lets us discard unwanted rows of a data frame (or keep the ones we do want). We need something like that, but for lists.

`purrr` has functions to do that, called `keep` and `discard`. See the [help file here](https://purrr.tidyverse.org/reference/keep.html).

`discard` works as follows: For a list object `x` and a function `logical_condition` that returns `TRUE` or `FALSE` for each element of the list,
discard(x, .p = logical_condition)
```

returns a list **dropping all elements for which `my_function` gives `TRUE`.**

`logical_condition` can be the name of a function, or an 'anonymous' function like the ones we saw in the strings lectures using the `~`.

#### Dummy example with outside function

```{r}
dumb_list <- list(a = c(1, 2), b = c(1, 2, 3, 4), d = c(-1, -2, 3))

dumb_f <- function(y){
  max(y) > 2
}


discard(dumb_list, .p = dumb_f)
```

## Question instructions

Using `discard` or `keep` and the function you made in Q2, take the output from Q1 and return a list containing removing the cases where the `santa` is the same as the `giftee`.

Save over your result from Q1 and display it.


```{r}
pairs <- discard(allPairs, .p = my_function)
pairs
```
#### Why not use a loop?
At first thought, the most intuitive way to solve this might be a loop.

Say you write a for loop that does the following

- iterates over the number of elements in the list output from Q1
- in each element
  - if `santa` is the same as `giftee`, replace the element with `NULL`. This removes it from the list.
  - otherwise, do not change anything

This would look like this, if the output from Q1 is called `xmas`

```{r,eval = FALSE}
for (i in 1:length(allPairs))
  if(allPairs[[i]][1]==allPairs[[i]][2])
    allPairs[[i]] <- NULL
```

But you are changing the length of the `xmas` list as you remove items! So when your iterator gets to index 16, the last index, your list will no longer have a 16th index because one has been removed. 

Try it and see. But don't forget to re-generate your output from Q1 before answering Q2, else the loop above will mess it up.

# Q4
**Repeat Q3** but instead of using an outside function in your `keep`/`discard` statement, use an anonymous function.

In other words, 

- **Rerun** the code from Q1
- **Using `discard` or `keep` and an anonymous function**, take the output from Q1 and return a list containing removing the cases where the `santa` is the same as the `giftee`. 
- **Save** over your result from Q1 and display it.

```{r}
pairs <- discard(allPairs, .p = function(x){
  if (x[1] == x[2])
    TRUE
  else 
    FALSE})
pairs
```


#### Dummy example of `discard` with anonymous function

```{r}
dumb_list <- list(a = c(1, 2), b = c(1, 2, 3, 4), d = c(-1, -2, 3))
```

**This doesn't work**
```{r, eval =F}
discard(dumb_list, .p = max > 2)
```

Here, use `.x` for an arbitrary element of your list `dumb_list`. The `~` tells it you are giving it an anonymous function.
```{r}
discard(dumb_list, .p = ~ max(.x) > 2)
```

#### Another version

You could also define the function within `discard` itself.

```{r}
discard(dumb_list, .p = function(x){
  max(x) > 2
})
```
# Q5 + 6
This is a double question since it might take more time.

**Write a function called `elf`** that does the following

i) takes as input a list with the same structure as the one output in Q3 or Q4
ii) randomly selects one element of the list, saving it as `choice`
iii) removes all elements of the input for which `santa` is the same as the one in `choice` OR `giftee` is the same as the one for `choice`
iv) returns a list with two names elements:
  - `choice` giving the chosen pair, result from step ii
  - `rest` giving the result of step iii
  
```{r}
elf <- function(x){
  choice <- sample(x,1)
  rest <- discard(x, .p = function(x){
  if (x[1] == choice[[1]][1]){
    TRUE}
  else if (x[2] == choice[[1]][2]){
    TRUE}
  else
    FALSE})
  output <- list("choice" = choice,"rest" = rest)
  return(output)}

elf(pairs)
```


#### Hints
`sample` works on lists as well as vectors. It will randomly select an entire element of the list. 

For example
```{r}
sample(list(c(1, 2), c("b", "a")), 1)
```
**Notice the output is a list itself**, of length one.

For part iii), use `discard` or `keep` again with an appropriate argument for `.p`.


# Q7 + 8
**Write a for loop** that does the following

- runs one iteration of your `elf` function using the `rest` part of the output of `elf` from the *previous* iteration and saves the entire output of `elf` for future use

**Show the output,** which will be a list in which each element is also a list with two elements, `choice` and `rest` as output by the `elf` function.

```{r}
output <- vector("list",length = 4)
output[[1]] <- elf(pairs)
nextpick <- output[[1]][[2]]
names(output) <- c("Pick 1","Pick 2","Pick 3","Pick 4")

for (i in 2:4){
  output[[i]] <- elf(nextpick)
  nextpick <- output[[i]][[2]]
}

output
```

#### Hint
The instruction might be a little unclear. I will write a dummy loop that does something analogous but for a different situation.

Say I have a vector `x = c(1, 2, 3, 4)` and I want to get the cumulative sum using a loop. (For real, though, don't do it this way. There is a function for that. This is just for demonstration).

I could do it this way

```{r}
x <- 1:4
out <- vector('numeric', length(x))

out <- x[1]

for (i in 2:length(x))
  out[i] <- x[i] + out[i-1]

out
```


# Q9
**Extract just the `choices` from the output from the previous question.** Display the result.

**This is your secret santa list of pairings!**

Do this any way you like: a loop or `map` with an anonymous function will work here.

```{r}
for (i in 1:4){
  print(output[[i]]$choice)}
```