---
title: "Programming Part B Tutorial"
output: html_notebook
---


```{r}
df = data.frame(team_one = 1:8, team_two = 9:16)
apply(df, 1, sample, size = 1)
```


```{r}
df = data.frame(team_one = 1:8, team_two = 9:16)

win = apply(df, 1, sample, size = 1)
matrix(win, ncol = 2, byrow = TRUE)
```
```{r}
dumb_list <- list(a = c(1, 2), b = c(1, 2, 3, 4), d = c(-1, -2, 3))

dumb_f <- function(y){
  max(y) > 2
}


discard(dumb_list, .p = dumb_f)
```

```{r}
discard(allPairs, .p = my_function)
```
```{r}
for (i in 1:length(xmas))
  if(xmas[[i]][1]==xmas[[i]][2])
    xmas[[i]] <- NULL
```


```{r}
for (i in 1:length(allPairs))
  if(allPairs[[i]][1]==allPairs[[i]][2])
    allPairs[[i]] <- NULL
```

```{r}
pairs <- discard(allPairs, .p = function(x){
  if (x[1] == x[2])
    TRUE
  else 
    FALSE})
pairs
```
my_function <- function(x){
  if (x[1] == x[2])
    TRUE
  else 
    FALSE
}
allPairs[[1]]
my_function(allPairs[[1]])

allPairs[[2]]
my_function(allPairs[[2]])

```{r}
newlist <- list()
for (i in 1){
  newlist <- elf(pairs)}

for (i in rest){
  choice <- sample(i,1)
  rest <- discard(i, .p = function(x){
  if (x[1] == choice[[1]][1]){
    TRUE}
  else if (x[2] == choice[[1]][2]){
    TRUE}
  else
    FALSE})
}


elf <- function(x){
  choice <- sample(x,1)
  rest <- discard(x, .p = function(x){
  if (x[1] == choice[[1]][1]){
    TRUE}
  else if (x[2] == choice[[1]][2]){
    TRUE}
  else
    FALSE})
  output <- list("choice" = choice,"rest" = rest)
  return(output)}
choice
rest
elf(pairs)
```
```{r}
newlist <- list()

for (i in 1:length(pairs)){
  choice <- pairs[i]
  rest <- discard(pairs, .p = function(x){
  if (x[1] == choice[[1]][1]){
    TRUE}
  else if (x[2] == choice[[1]][2]){
    TRUE}
  else
    FALSE})
  newlist <- list("choice" = choice,"rest" = rest)
  print(newlist)}

newlist
  
```
#### Hint
The instruction might be a little unclear. I will write a dummy loop that does something analogous but for a different situation.

Say I have a vector `x = c(1, 2, 3, 4)` and I want to get the cumulative sum using a loop. (For real, though, don't do it this way. There is a function for that. This is just for demonstration).

I could do it this way
```{r}
x <- 1:4
out <- vector('numeric', length(x))
out
out <- x[1]
out
for (i in 2:length(x))
  out[i] <- x[i] + out[i-1]

out
```

```{r}
output <- vector("list",length = 4)
output[[1]] <- elf(pairs)
nextpick <- output[[1]][[2]]
names(output) <- c("Pick 1","Pick 2","Pick 3","Pick 4")

for (i in 2:4){
  output[[i]] <- elf(nextpick)
  nextpick <- output[[i]][[2]]
}

for (i in 1:4){
  print(output[[i]]$choice)}


```

